---
title: "Analysis_Tcf21_CM_new"
author: "Giulia"
date: "24/7/2022"
output: html_document
---

---
title: "Analysis_Tcf21_CM_new"
author: "Giulia Boezio"
date: "24/7/2022"
output: html_document
---

# Load libraries & setup

```{r message=FALSE}
library(tidyverse)
library(readxl)
library(clusterProfiler)
library(UpSetR)
library(viridis)
library(org.Dr.eg.db)
library(biomaRt)
library(writexl)
library(ReactomePA)
```

```{r message=FALSE, echo = FALSE}
dir_save = "/Users/boeziog/R_analyses_epi/data"
dir_save_plots = "/Users/boeziog/R_analyses_epi/plots"
file_input = "/Users/boeziog/R_analyses_epi/data/SHORTdst141_RNAseq_Giulia_01_21_rnaseq.xlsx"
```

# Load data

```{r}
tot_df = read_xlsx(file_input, sheet = "matrix",
                   .name_repair = function(x) gsub("\\s+", "_", x)) %>% 
  dplyr::select(Ensembl_gene_id, Ensembl_gene, starts_with("CM"), starts_with("Epi"))
tot_df

CM = read_xlsx(file_input, sheet = "CM-Mut_vs_CM-WT",
                   .name_repair = function(x) gsub("\\s+", "_", x)) %>%
  # dplyr::rename(log2FoldChange = log2FoldChange_CM-Mut/CM-WT) %>%
  dplyr::select(Ensembl_gene_id, Ensembl_gene, log2FoldChange, padj, pvalue)
CM

Epi = read_xlsx(file_input, sheet = "Epi-Mut_vs_Epi-WT",
                   .name_repair = function(x) gsub("\\s+", "_", x)) %>%
  # dplyr::rename(log2FoldChange = "log2FoldChange_Epi-Mut/Epi-WT") %>% 
  dplyr::select(Ensembl_gene_id, Ensembl_gene, log2FoldChange, padj, pvalue)
Epi


WT = read_xlsx(file_input, sheet = "Epi-WT_vs_CM-WT",
                   .name_repair = function(x) gsub("\\s+", "_", x)) %>%
#   dplyr::rename(log2FoldChange = "log2FoldChange_Epi-WT/CM-WT") %>% 
  dplyr::select(Ensembl_gene_id, Ensembl_gene, log2FoldChange, padj, pvalue)
WT

```

# Select DEGs

```{r}
min_Log2 = 0.585
max_pvalue = 0.05


Epi_spec = WT %>% filter(log2FoldChange > min_Log2 & pvalue < max_pvalue)
CM_spec = WT %>% filter(log2FoldChange < -min_Log2 & pvalue < max_pvalue)
Epi_Up = Epi %>% filter(log2FoldChange > min_Log2 & pvalue < max_pvalue)
Epi_Down = Epi %>% filter(log2FoldChange < -min_Log2 & pvalue < max_pvalue)
CM_Up = CM %>% filter(log2FoldChange > min_Log2 & pvalue < max_pvalue)
CM_Down = CM %>% filter(log2FoldChange < -min_Log2 & pvalue < max_pvalue)
```

### Excel
```{r}
write_xlsx(CM_Up,
           path = file.path(dir_save, paste0("DEG_", "CM_Up",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             ".xlsx")))

write_xlsx(CM_Down,
           path = file.path(dir_save, paste0("DEG_", "CM_Down",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             ".xlsx")))
write_xlsx(Epi_Up,
           path = file.path(dir_save, paste0("DEG_", "Epi_Up",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             ".xlsx")))

write_xlsx(Epi_Down,
           path = file.path(dir_save, paste0("DEG_", "Epi_Down",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             ".xlsx")))
write_xlsx(Epi_spec,
           path = file.path(dir_save, paste0("DEG_", "Epi_Spec",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             ".xlsx")))
write_xlsx(CM_spec,
           path = file.path(dir_save, paste0("DEG_", "CM_spec",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             ".xlsx")))
```

## Upset plot

```{r}
UpInput = list(
      WT_Epi_specific = Epi_spec %>% pull(Ensembl_gene_id),
               WT_CM_specific = CM_spec %>% pull(Ensembl_gene_id),
               Epi_Up = Epi_Up %>% pull(Ensembl_gene_id),
               Epi_Down = Epi_Down %>% pull(Ensembl_gene_id),
               CM_Up = CM_Up %>% pull(Ensembl_gene_id),
               CM_Down = CM_Down %>% pull(Ensembl_gene_id))

pdf(file.path(dir_save_plots, paste0("Upset_", "Log2FC_", min_Log2, "_pvalue_", max_pvalue, ".pdf")),  onefile=FALSE)

upset(fromList(UpInput),
      sets = rev(c(
            "WT_Epi_specific",
                   "WT_CM_specific",
                   "Epi_Up",
                   "Epi_Down",
                   "CM_Up",
                   "CM_Down")),
      mainbar.y.label = "DGEs", sets.x.label = "Genes per set", 
      keep.order = TRUE, order.by = c("freq"),
      text.scale = c(1.75, 1.75, 1.5, 1.5, 2, 1.2))
      # c(1.5, 2, 1.5, 1.3, 1.7, 1.8) c(intersection size title, intersection size tick labels, set size title, set size tick labels, set names, numbers above bars)
dev.off()

```

#GO 
## Prepare data 
### Get EntrezID

```{r}
ensembl = useEnsembl(biomart="ensembl", dataset="drerio_gene_ensembl")

tot_genes = tot_df %>% pull(Ensembl_gene_id)

tot_genes = getBM(attributes = c('entrezgene_id', 'ensembl_gene_id'),
      filters = 'ensembl_gene_id',
      values = tot_genes,
      mart = ensembl, useCache = FALSE)

dim(tot_df)
dim(tot_genes)

# there are genes with multiple entrezID matches. Remove them
multi_matchers = tot_genes %>% group_by(ensembl_gene_id) %>% 
  summarize(n = n()) %>% filter(n > 1) %>% arrange(by = desc(n))
tot_genes = anti_join(tot_genes, multi_matchers)
dim(tot_genes)
```

### Convert each subset to EntrezID

```{r}
Epi_spec = Epi_spec %>% left_join(tot_genes, by = c("Ensembl_gene_id" = "ensembl_gene_id")) %>%
  mutate(set = "Epi_specific")
CM_spec = CM_spec %>% left_join(tot_genes, by = c("Ensembl_gene_id" = "ensembl_gene_id")) %>%
  mutate(set = "CM_specific")
Epi_Up = Epi_Up %>% left_join(tot_genes, by = c("Ensembl_gene_id" = "ensembl_gene_id")) %>% 
  mutate(set = "Epi_Up")
Epi_Down = Epi_Down %>% left_join(tot_genes, by = c("Ensembl_gene_id" = "ensembl_gene_id")) %>% 
  mutate(set = "Epi_Down")
CM_Up = CM_Up %>% left_join(tot_genes, by = c("Ensembl_gene_id" = "ensembl_gene_id")) %>% 
  mutate(set = "CM_Up")
CM_Down = CM_Down %>% left_join(tot_genes, by = c("Ensembl_gene_id" = "ensembl_gene_id")) %>% 
  mutate(set = "CM_Down")

paste("Epi_spec missing match:",
      Epi_spec %>% filter(is.na(entrezgene_id)) %>% pull(entrezgene_id) %>%  length(),
      " Remaining :", dim(Epi_spec)[1], "genes")

paste("CM_spec missing match:",
      CM_spec %>% filter(is.na(entrezgene_id)) %>% pull(entrezgene_id) %>%  length(),
      " Remaining :", dim(CM_spec)[1], "genes")

paste("Epi_Up missing match:",
      Epi_Up %>% filter(is.na(entrezgene_id)) %>% pull(entrezgene_id) %>%  length(),
      " Remaining :", dim(Epi_Up)[1], "genes")

paste("Epi_Down missing match:",
      Epi_Down %>% filter(is.na(entrezgene_id)) %>% pull(entrezgene_id) %>%  length(),
      " Remaining :", dim(Epi_Down)[1], "genes")

paste("CM_Up missing match:",
      CM_Up %>% filter(is.na(entrezgene_id)) %>% pull(entrezgene_id) %>%  length(),
      " Remaining :", dim(CM_Up)[1], "genes")

paste("CM_Down missing match:",
      CM_Down %>% filter(is.na(entrezgene_id)) %>% pull(entrezgene_id) %>%  length(),
      " Remaining :", dim(CM_Down)[1], "genes")
```

### Make cluster profiler df

```{r}
cProf_data = bind_rows(
      Epi_spec, CM_spec,
      Epi_Up, Epi_Down, CM_Up, CM_Down) %>%
  filter(!is.na(entrezgene_id))
cProf_data
```

## Define compareCluster parameters

```{r}
fun           = "enrichGO"
universe      = tot_genes %>% pull(entrezgene_id)
OrgDb         = org.Dr.eg.db
keyType       = "ENTREZID"
pAdjustMethod = "BH"
minGSSize     = 10
maxGSSize     = 500
pvalueCutoff  = 0.05
qvalueCutoff  = 0.05
```

## Set simplification & colour palette parameters
```{r}
#Set semplification and plotting parameters (raggruppa categorie ridondanti, cutoff basso raggruppa (1, ognuno per conto suo, 0= tutto insieme).  
cutoff     = 0.7
by         = "p.adjust"
select_fun = min

# viridis parameters
option    = 'viridis'
direction = -1 # reverse color map order
```

## Run ClusterProfiler

### BP

```{r}
ont = "BP"
Epi_comparisons = cProf_data %>% filter(set == "Epi_Up" | set == "Epi_Down")
CM_comparisons = cProf_data %>% filter(set == "CM_Up" | set == "CM_Down")
WT_comparisons = cProf_data %>% filter(set == "Epi_specific" | set == "CM_specific")

Epi_BP <- compareCluster(entrezgene_id~set, data=Epi_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize, 
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

CM_BP <- compareCluster(entrezgene_id~set, data=CM_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize, 
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

WT_BP <- compareCluster(entrezgene_id~set, data=WT_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize,
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)
```

#### Simplification
```{r}
#Simplify GO
Epi_simpler_BP = clusterProfiler::simplify(Epi_BP, cutoff=cutoff, by=by, select_fun=select_fun)
CM_simpler_BP = clusterProfiler::simplify(CM_BP, cutoff=cutoff, by=by, select_fun=select_fun)
WT_simpler_BP = clusterProfiler::simplify(WT_BP, cutoff=cutoff, by=by, select_fun=select_fun)
```

#### Excel

```{r}
ont = "BP"
#save excel
write_xlsx(Epi_simpler_BP@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "Epi_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
write_xlsx(CM_simpler_BP@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "CM_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
write_xlsx(WT_simpler_BP@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "WT_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
```

#### Plot

```{r fig.height=3.5, fig.width=4}
#Plotting

# dotplot parameters
ont = "BP"
showCategory = 10
title        = "Biological Processes"
font_size    = 10

# shorten names
Epi_simpler_BP@compareClusterResult$Description = Epi_simpler_BP@compareClusterResult$Description %>% 
  gsub("system", "sys.",.) %>%
  gsub("development", "dev.",.) %>%
  gsub("organization", "org.",.) %>%
  gsub("membrane", "memb.",.) %>%
  gsub("projection", "proj.",.) %>%
  gsub("morphogenesis", "morph.",.) %>%
  gsub("regulation", "reg.",.) %>%
  gsub("proliferation", "prolif.",.) %>%
  gsub("negative", "neg.",.) %>%
  gsub("macromolecule", "macrom.",.) %>%
  gsub("biosynthetic", "biosyn.",.) %>%
  gsub("process", "proc.",.) %>%
  gsub("maintenance", "maint.",.) %>%
  gsub("positive", "pos.",.) %>%
  gsub(" of ", " ",.) %>%
  gsub("metabolic", "metab.",.) %>%
  gsub("establishment", "est.",.) %>%
  gsub("phosphorylation", "phos.",.) %>%
  gsub("transcription", "trxn.",.) %>%
  gsub("templated", "templ.",.) %>% 
  gsub("nucleobase", "nucleob.",.) %>%
  gsub("templated", "templ.",.) %>%
  gsub("cellullar", "cell.",.) %>%
  gsub("component", "comp.",.) %>%
  gsub("modification", "mod",.) %>%
  gsub("transport", "trans.",.) %>%
  gsub("protein", "prot.",.) %>%
  gsub("differentiation", "diff.",.) %>%
  gsub("containing", "cont.",.) %>%
  gsub("response", "resp.",.) %>%
  gsub("commitment", "comm.",.) %>%
  gsub("structure", "str.",.) %>%
  gsub("trandsuction", "transd.",.) %>%
  gsub("assembly", "ass.",.) %>%
  gsub("compound", "comp.",.) %>%
  gsub("specification", "spec.",.)
  #gsub("", "",.) %>%
  
CM_simpler_BP@compareClusterResult$Description = CM_simpler_BP@compareClusterResult$Description %>% 
  gsub("system", "sys.",.) %>%
  gsub("development", "dev.",.) %>%
  gsub("organization", "org.",.) %>%
  gsub("membrane", "memb.",.) %>%
  gsub("projection", "proj.",.) %>%
  gsub("morphogenesis", "morph.",.) %>%
  gsub("regulation", "reg.",.) %>%
  gsub("proliferation", "prolif.",.) %>%
  gsub("negative", "neg.",.) %>%
  gsub("macromolecule", "macrom.",.) %>%
  gsub("biosynthetic", "biosyn.",.) %>%
  gsub("process", "proc.",.) %>%
  gsub("maintenance", "maint.",.) %>%
  gsub("positive", "pos.",.) %>%
  gsub(" of ", " ",.) %>%
  gsub("metabolic", "metab.",.) %>%
  gsub("establishment", "est.",.) %>%
  gsub("phosphorylation", "phos.",.) %>%
  gsub("transcription", "trxn.",.) %>%
  gsub("templated", "templ.",.) %>% 
  gsub("nucleobase", "nucleob.",.) %>%
  gsub("templated", "templ.",.) %>%
  gsub("cellullar", "cell.",.) %>%
  gsub("component", "comp.",.) %>%
  gsub("modification", "mod",.) %>%
  gsub("transport", "trans.",.) %>%
  gsub("protein", "prot.",.) %>%
  gsub("differentiation", "diff.",.) %>%
  gsub("containing", "cont.",.) %>%
  gsub("response", "resp.",.) %>%
  gsub("commitment", "comm.",.) %>%
  gsub("structure", "str.",.) %>%
  gsub("trandsuction", "transd.",.) %>%
  gsub("assembly", "ass.",.) %>%
  gsub("compound", "comp.",.) %>%
  gsub("specification", "spec.",.)

Epi_simpler_BP@compareClusterResult$Description = Epi_simpler_BP@compareClusterResult$Description %>%
  gsub("system", "sys.",.) %>%
  gsub("development", "dev.",.) %>%
  gsub("organization", "org.",.) %>%
  gsub("membrane", "memb.",.) %>%
  gsub("projection", "proj.",.) %>%
  gsub("morphogenesis", "morph.",.) %>%
  gsub("regulation", "reg.",.) %>%
  gsub("proliferation", "prolif.",.) %>%
  gsub("negative", "neg.",.) %>%
  gsub("macromolecule", "macrom.",.) %>%
  gsub("biosynthetic", "biosyn.",.) %>%
  gsub("process", "proc.",.) %>%
  gsub("maintenance", "maint.",.) %>%
  gsub("positive", "pos.",.) %>%
  gsub(" of ", " ",.) %>%
  gsub("metabolic", "metab.",.) %>%
  gsub("establishment", "est.",.) %>%
  gsub("phosphorylation", "phos.",.) %>%
  gsub("transcription", "trxn.",.) %>%
  gsub("templated", "templ.",.) %>%
  gsub("nucleobase", "nucleob.",.) %>%
  gsub("templated", "templ.",.) %>%
  gsub("cellullar", "cell.",.) %>%
  gsub("component", "comp.",.) %>%
  gsub("modification", "mod",.) %>%
  gsub("transport", "trans.",.) %>%
  gsub("protein", "prot.",.) %>%
  gsub("differentiation", "diff.",.) %>%
  gsub("containing", "cont.",.) %>%
  gsub("response", "resp.",.) %>%
  gsub("commitment", "comm.",.) %>%
  gsub("structure", "str.",.) %>%
  gsub("trandsuction", "transd.",.) %>%
  gsub("assembly", "ass.",.) %>%
  gsub("compound", "comp.",.) %>%
  gsub("phosphate", "phosp.",.) %>%
      gsub("mitochondrion", "mito.",.) %>%
      gsub("mitochondrial", "mito.",.) %>%
      gsub("localization", "local.",.) %>%
  gsub("specification", "spec.",.)

# Plot
dotplot(Epi_simpler_BP, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
ggsave(file.path(dir_save_plots, paste0("GO_", "Epi_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")))

dotplot(CM_simpler_BP, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
ggsave(file.path(dir_save_plots, paste0("GO_", "CM_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")))

dotplot(WT_simpler_BP, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
# Save
ggsave(file.path(dir_save_plots, paste0("GO_", "WT_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")))
```

### MF

```{r}
# clusterProfiler
ont          = "MF"

Epi_MF <- compareCluster(entrezgene_id~set, data=Epi_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize, 
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

CM_MF <- compareCluster(entrezgene_id~set, data=CM_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize,
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

WT_MF <- compareCluster(entrezgene_id~set, data=WT_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize,
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)
```

#### Simplification

```{r}
#Set semplification parameters
cutoff     = 0.7
by         = "p.adjust"
select_fun = min

Epi_simpler_MF = clusterProfiler::simplify(Epi_MF, cutoff=cutoff, by=by, select_fun=select_fun)
CM_simpler_MF = clusterProfiler::simplify(CM_MF, cutoff=cutoff, by=by, select_fun=select_fun)
WT_simpler_MF = clusterProfiler::simplify(WT_MF, cutoff=cutoff, by=by, select_fun=select_fun)
```

#### Excel

```{r}
ont          = "MF"
write_xlsx(Epi_simpler_MF@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "Epi_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))

write_xlsx(CM_simpler_MF@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "CM_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                                "_minGSSize_", minGSSize,
                                                "_maxGSSize_", maxGSSize,
                                                "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                                ".xlsx")))
write_xlsx(WT_simpler_MF@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "WT_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
```

#### Plot

```{r fig.height=3.5, fig.width=4}

# dotplot parameters
showCategory = 10
ont          = "MF"
title        = "Molecular Functions"
font_size    = 9

dotplot(Epi_simpler_MF, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
ggsave(file.path(dir_save_plots, paste0("GO_", "Epi_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")), 
       height = 5, width = 5)

dotplot(CM_simpler_MF, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
ggsave(file.path(dir_save_plots, paste0("GO_", "CM_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                                "_minGSSize_", minGSSize,
                                                "_maxGSSize_", maxGSSize,
                                                "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                                "_dotplot.pdf")),
                              height = 5.5, width = 5)

dotplot(WT_simpler_MF, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
# Save
ggsave(file.path(dir_save_plots, paste0("GO_", "WT_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")),
                        height = 7, width = 5)
```

### CC

```{r}
# clusterProfiler
ont          = "CC"

title        = "Cellular Compartment"

Epi_CC <- compareCluster(entrezgene_id~set, data=Epi_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize,
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

CM_CC <- compareCluster(entrezgene_id~set, data=CM_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize,
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

WT_CC <- compareCluster(entrezgene_id~set, data=WT_comparisons, readable = TRUE,
                              fun = fun, universe = universe, OrgDb = OrgDb,
                              ont = ont, keyType = keyType, pAdjustMethod = pAdjustMethod,
                              minGSSize = minGSSize,maxGSSize = maxGSSize,
                              pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)
```

#### Simplification

```{r}
#Set semplification parameters
cutoff     = 0.7
by         = "p.adjust"
select_fun = min

Epi_simpler_CC = clusterProfiler::simplify(Epi_CC, cutoff=cutoff, by=by, select_fun=select_fun)
CM_simpler_CC = clusterProfiler::simplify(CM_CC, cutoff=cutoff, by=by, select_fun=select_fun)
WT_simpler_CC = clusterProfiler::simplify(WT_CC, cutoff=cutoff, by=by, select_fun=select_fun)
```

#### Excel

```{r}
ont          = "CC"
write_xlsx(Epi_simpler_CC@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "Epi_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
"_minGSSize_", minGSSize,
"_maxGSSize_", maxGSSize,
"_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
".xlsx")))

write_xlsx(CM_simpler_CC@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "CM_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))

write_xlsx(WT_simpler_CC@compareClusterResult,
           path = file.path(dir_save, paste0("GO_", "WT_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
```

#### Plot

```{r fig.height=6, fig.width= 5}
ont          = "CC"
dotplot(Epi_simpler_CC, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
ggsave(file.path(dir_save_plots, paste0("GO_", "Epi_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                          "_minGSSize_", minGSSize,
                                          "_maxGSSize_", maxGSSize,
                                          "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                          "_dotplot.pdf")),
                                                height = 5, width = 5)

dotplot(CM_simpler_CC, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
ggsave(file.path(dir_save_plots, paste0("GO_", "CM_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")),
                                    height = 4.5, width = 4.3)

dotplot(WT_simpler_CC, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    # legend.position="bottom", legend.direction = "horizontal", legend.box = "vertical",
                          )
# Save
ggsave(file.path(dir_save_plots, paste0("GO_", "WT_", ont,
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")),
       height = 6, width = 4.5)
```

## ReactomePA

```{r}
fun = "enrichPathway"
title        = "Reactome Pathways Enriched"
minGSSize     = 10
maxGSSize     = 1000
pvalueCutoff  = 0.05
qvalueCutoff  = 0.05
```

###Epicardium

```{r}
##########################################
Epi_Path <- compareCluster(entrezgene_id~set, data=Epi_comparisons,
                          organism = "zebrafish", readable = TRUE,
                          fun = fun, universe = universe,
                          pAdjustMethod = pAdjustMethod,
                          minGSSize = minGSSize, maxGSSize = maxGSSize,
                          pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

Epi_Path@compareClusterResult$Description = Epi_Path@compareClusterResult$Description %>%
  gsub("Respiratory", "Resp.",.) %>%
  gsub("electron", "e-",.) %>%
  gsub("transport", "trsp.",.) %>%
  gsub("synthesis", "synth.",.) %>%
  gsub("chemiosmotic", "chosm.",.) %>%
  gsub("coupling", "cpl.",.) %>%
  gsub("production", "prod.",.) %>%
  gsub("uncoupling", "uncpl.",.) %>%
  gsub("protein", "prot.",.)
  # gsub("Respiratory", "Resp.",.) %>%
```

####Excel

```{r}
write_xlsx(Epi_Path@compareClusterResult,
           path = file.path(dir_save, paste0("ReactomePA_", "Epi_",
                                             "log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
```

####Plot

```{r fig.height=1.5, fig.width=3.5}
dotplot(Epi_Path, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    )

ggsave(file.path(dir_save_plots, paste0("ReactomePA_", "Epi_",
                                             "log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")),
       height = 3)
```

###CM

```{r}
##########################################
CM_Path <- compareCluster(entrezgene_id~set, data=CM_comparisons,
                          organism = "zebrafish", readable = TRUE,
                          fun = fun, universe = universe,
                          pAdjustMethod = pAdjustMethod,
                          minGSSize = minGSSize, maxGSSize = maxGSSize,
                          pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

CM_Path@compareClusterResult$Description = CM_Path@compareClusterResult$Description %>% 
  gsub("Respiratory", "Resp.",.) %>%
  gsub("electron", "e-",.) %>%
  gsub("transport", "trsp.",.) %>%
  gsub("synthesis", "synth.",.) %>%
  gsub("chemiosmotic", "chosm.",.) %>%
  gsub("coupling", "cpl.",.) %>%
  gsub("production", "prod.",.) %>%
  gsub("uncoupling", "uncpl.",.) %>%
  gsub("protein", "prot.",.)
  # gsub("Respiratory", "Resp.",.) %>%
```

####Excel

```{r}
write_xlsx(CM_Path@compareClusterResult,
           path = file.path(dir_save, paste0("ReactomePA_", "CM_",
                                             "log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
```

####Plot

```{r fig.height=1.5, fig.width=3.5}

# dotplot parameters
showCategory = 10
title        = "Reactome Pathways Enriched"
font_size    = 10


dotplot(CM_Path, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    )

ggsave(file.path(dir_save_plots, paste0("ReactomePA_", "CM_",
                                             "log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,"_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")),
      height = 6, width = 6)
```

###WT

```{r}
################################################
WT_Path <- compareCluster(entrezgene_id~set, data=WT_comparisons,
                          organism = "zebrafish", readable = TRUE,
                          fun = fun, universe = universe,
                          pAdjustMethod = pAdjustMethod,
                          minGSSize = minGSSize, maxGSSize = maxGSSize,
                          pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)

WT_Path@compareClusterResult$Description = WT_Path@compareClusterResult$Description %>%
  gsub("Respiratory", "Resp.",.) %>%
  gsub("electron", "e-",.) %>%
  gsub("transport", "trsp.",.) %>%
  gsub("synthesis", "synth.",.) %>%
  gsub("chemiosmotic", "chosm.",.) %>%
  gsub("coupling", "cpl.",.) %>%
  gsub("production", "prod.",.) %>%
  gsub("uncoupling", "uncpl.",.) %>%
  gsub("protein", "prot.",.)
  # gsub("Respiratory", "Resp.",.) %>%
```

####Excel

```{r}

write_xlsx(WT_Path@compareClusterResult,
           path = file.path(dir_save, paste0("ReactomePA_", "WT_",
                                             "log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             ".xlsx")))
```

####Plot

```{r fig.height=6, fig.width=4}

# dotplot parameters
showCategory = 10
title        = "Reactome Pathways Enriched"
font_size    = 10

dotplot(WT_Path, showCategory = showCategory, title = title, font.size = font_size) +
                    scale_color_viridis(option=option, direction = direction) +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.0),
                    )

ggsave(file.path(dir_save_plots, paste0("ReactomePA_", "WT_",
                                             "log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue,"_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_pval_", pvalueCutoff, "_qval_", qvalueCutoff,
                                             "_dotplot.pdf")),
       height = 8.5, width = 6)
```

```{r}
dim(Epi_comparisons)
```

# Session Info

```{r}
sessionInfo()
```
