---
title: "GSEA"
author: "Giulia"
date: '2022-07-27'
output: html_document
---
---
title: "Gene Set Enrichment Analysis with ClusterProfiler"
author: "Mohammed Khalfan"
date: "5/19/2019"
output:
  html_document:
    df_print: paged
df_print: paged
---

This R Notebook describes the implementation of gene set enrichment analysis (GSEA) using the clusterProfiler package. For more information please see the full documentation here: https://bioconductor.org/packages/release/bioc/vignettes/clusterProfiler/inst/doc/clusterProfiler.html


# Install and load required packages
```{r, message=F, warning=F}
#BiocManager::install("clusterProfiler", version = "3.8")
#BiocManager::install("pathview")
#BiocManager::install("enrichplot")
library(clusterProfiler)
library(enrichplot)
library(ggplot2)
library(tidyverse)
library(readxl)
library(viridis)
library(org.Dr.eg.db)
library(biomaRt)
library(writexl)
# devtools::install_github("nicolash2/gggsea")
# library(gggsea)
```
# Directories
```{r message=FALSE, echo = FALSE}
dir_save = "/Users/boeziog/R_analyses_epi/data"
dir_save_plots = "/Users/boeziog/R_analyses_epi/plots"
file_input = "/Users/boeziog/R_analyses_epi/data/SHORTdst141_RNAseq_Giulia_01_21_rnaseq.xlsx"
CM_all = "/Users/boeziog/R_analyses_epi/data/CM_all.xlsx"
Epi_all = "/Users/boeziog/R_analyses_epi/data/Epi_all.xlsx"
Spec_all = "/Users/boeziog/R_analyses_epi/data/Specific_all.xlsx"
```

#Prepare Input
```{r}
# tot_df = read_xlsx(file_input, sheet = "matrix",
#                    .name_repair = function(x) gsub("\\s+", "_", x)) %>% 
#   dplyr::select(Ensembl_gene_id, Ensembl_gene, starts_with("CM"), starts_with("Epi"))
# tot_df


#CM
CM = read_xlsx(CM_all,
                   .name_repair = function(x) gsub("\\s+", "_", x))%>%
  # dplyr::rename(log2FoldChange = log2FoldChange_CM-Mut/CM-WT) %>%
dplyr::select(ensembl_gene_id,log2FoldChange, pvalue)%>%
      filter(pvalue < 0.05) %>%
      as.tibble()%>%
      # column_to_rownames(var = "ensembl_gene_id")%>%
arrange(desc(log2FoldChange))
CM

CM_gene_list = CM %>% 
      pull(log2FoldChange)

names(CM_gene_list) = CM %>% 
      pull(ensembl_gene_id) 

head(CM_gene_list)

#EPI

Epi = read_xlsx(Epi_all,
                   .name_repair = function(x) gsub("\\s+", "_", x))%>%
  # dplyr::rename(log2FoldChange = log2FoldChange_CM-Mut/CM-WT) %>%
dplyr::select(ensembl_gene_id,log2FoldChange, pvalue)%>%
      filter(pvalue < 0.05) %>%
      as.tibble()%>%
      # column_to_rownames(var = "ensembl_gene_id")%>%
arrange(desc(log2FoldChange))
Epi

Epi_gene_list = Epi %>% 
      pull(log2FoldChange)

names(Epi_gene_list) = Epi %>% 
      pull(ensembl_gene_id) 

head(Epi_gene_list)

#Cell type specific


Specific = read_xlsx(Spec_all,
                   .name_repair = function(x) gsub("\\s+", "_", x))%>%
  # dplyr::rename(log2FoldChange = log2FoldChange_CM-Mut/CM-WT) %>%
dplyr::select(ensembl_gene_id,log2FoldChange, pvalue)%>%
      filter(pvalue < 0.05) %>%
      as.tibble()%>%
      # column_to_rownames(var = "ensembl_gene_id")%>%
arrange(desc(log2FoldChange))
Specific

Spec_gene_list = Specific %>% 
      pull(log2FoldChange)

names(Spec_gene_list) = Specific %>% 
      pull(ensembl_gene_id) 

head(Spec_gene_list)

```


# Annotations
 I install and load the annotation "org.Dr.eg.db" below. See all annotations available here: http://bioconductor.org/packages/release/BiocViews.html#___OrgDb (there are 19 presently available). 

```{r, message=F, warning=F}
# SET THE DESIRED ORGANISM HERE
organism = "org.Dr.eg.db"
#BiocManager::install(organism, character.only = TRUE)
library(organism, character.only = TRUE)
```


## Gene Set Enrichment
Params:  
  
**keyType** This is the source of the annotation (gene ids). The options vary for each annotation.
  
Check which options are available with the `keytypes` command, for example `keytypes(org.Dm.eg.db)`. 
  
**ont** one of "BP", "MF", "CC" or "ALL"  
**nPerm** permutation numbers, the higher the number of permutations you set, the more accurate your results is, but it will also cost longer time for running permutation.  
**minGSSize** minimal size of each geneSet for analyzing.   
**maxGSSize** maximal size of genes annotated for testing.   
**pvalueCutoff** pvalue Cutoff.   
**pAdjustMethod** one of "holm", "hochberg", "hommel", "bonferroni", "BH", "BY", "fdr", "none" 


```{r}

nPerm = 100000 
minGSSize = 10 
maxGSSize = 300
pvalueCutoff = 0.05

gse <- gseGO(geneList=CM_gene_list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = nPerm, 
             minGSSize = minGSSize , 
             maxGSSize = maxGSSize, 
             pvalueCutoff = pvalueCutoff, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

gse_epi <- gseGO(geneList=Epi_gene_list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = nPerm, 
             minGSSize = minGSSize , 
             maxGSSize = maxGSSize, 
             pvalueCutoff = pvalueCutoff, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

gse_spec <- gseGO(geneList=Spec_gene_list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = nPerm, 
             minGSSize = minGSSize , 
             maxGSSize = maxGSSize, 
             pvalueCutoff = pvalueCutoff, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")



```

# Output
##Table of results
```{r}
head(gse, n = 20)

head(gse_epi, n = 20)

head(gse_spec, n = 20)

```

##Dotplot
```{r echo=TRUE, fig.width=14, fig.height=10}
require(DOSE)

#CM
dotplot(gse, showCategory= 10, split=".sign", font.size = 10, label_format = 30, title = "Gene Set Enrichment - GO terms CM") + facet_grid(.~.sign) + scale_color_viridis(option=option, direction = -1) 

ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "CM_",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_dotplot.pdf")), 
        height = 6, width = 10)

#Epi
dotplot(gse_epi, showCategory= 10, split=".sign", font.size = 10, label_format = 30, title = "Gene Set Enrichment - GO terms EpiCs") + facet_grid(.~.sign) + scale_color_viridis(option=option, direction = -1) 

ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "Epi_",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_dotplot.pdf")), 
        height = 6, width = 10)

#Specific

dotplot(gse_spec, showCategory= 10, split=".sign", font.size = 10, label_format = 30, title = "Gene Set Enrichment - GO terms EpiC vs CMs") + facet_grid(.~.sign) + scale_color_viridis(option=option, direction = -1) 

ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "Spec",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_dotplot.pdf")), 
        height = 6, width = 10)

```

## Ridgeplot
Helpful to interpret up/down-regulated pathways.
```{r fig.width=18, fig.height=12}
#CM
ridgeplot(gse,showCategory = 20,label_format = 25 ) + labs(x = "enrichment distribution") + scale_fill_viridis(option=option, direction = -1)
                   

ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "CM_",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_ridgeplot.pdf")), 
        height = 14, width = 25)

#Epi
ridgeplot(gse_epi) + labs(x = "enrichment distribution") + scale_fill_viridis(option=option, direction = -1)
                   
ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "Epi",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_ridgeplot.pdf")), 
        height = 20, width = 20)

#Specific
ridgeplot(gse_spec) + labs(x = "enrichment distribution") + scale_fill_viridis(option=option, direction = -1)
                   

ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "Spec",
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_ridgeplot.pdf")), 
        height = 12, width = 20)

```

## GSEA Plot  
Traditional method for visualizing GSEA result.  

Params:  
**Gene Set** Integer. Corresponds to gene set in the gse object. The first gene set is 1, second gene set is 2, etc. 

```{r fig.height=6}
# Use the `Gene Set` param for the index in the title, and as the value for geneSetId

#CM
geneSetID = 1
# gseaplot(gse, by = "runningScore", title = gse$Description[geneSetID], geneSetID = geneSetID,  color = "black", color.line = "green",color.vline = "red")

gseaplot2(
  gse,
  geneSetID = geneSetID,
  title = gse$Description[geneSetID],
  color = "blue",
  base_size = 13,
  rel_heights = c(2, 0.5, 1),
  subplots = 1:2,
  pvalue_table = FALSE,
  ES_geom = "line"
) 

ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "CM_",
                                             gse$Description[geneSetID], 
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_gseaplot.pdf")), 
        height = 6, width = 10)

#Epi
geneSetID = 1

gseaplot2(gse_epi,
  geneSetID = geneSetID,
  title = gse$Description[geneSetID],
  color = "blue",
  base_size = 13,
  rel_heights = c(2, 0.5, 1),
  subplots = 1:2,
  pvalue_table = FALSE,
  ES_geom = "line"
) 


ggsave(file.path(dir_save_plots, paste0("GSEA_GO_", "Epi_",
                                             gse$Description[geneSetID], 
                                             "_log2FC_", min_Log2,
                                             "_pvalue_", max_pvalue, 
                                             "_minGSSize_", minGSSize,
                                             "_maxGSSize_", maxGSSize,
                                             "_nPerm_", nPerm,
                                             "_pvalCutoff_", pvalueCutoff,
                                             "_gseaplot.pdf")), 
        height = 6, width = 10)
```
